---
title: "TB Hotspot clusters P:N"

subtitle: "Exploratory analysis of the neighbourhood level data"

author: |
  | McEwen Khundi
  |
  | London School of Hygiene and Tropical Medicine, London, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    df_print: paged
    theme: spacelab
    toc: true
---

<br>

## 1. Backgound

<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, message=FALSE}
library(tidyverse)    #for data manipulation
library(pmthemes)     #for ggplot themes
library(knitr)        #for tables
library(arsenal)
library(sf)
library(scales)
library(viridis)
library(patchwork)
library(flextable)

select <- dplyr::select
```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import}
#load(here::here("data/all_cases_sf.rda"))
#load(here::here("data/cluster_cases_sf.rda"))
load(here::here("data/clinics.rda"))
load(here::here("data/cnrs.rda"))
load(here::here("data/scale_clusters.rda"))
dat_scale <- readRDS(here::here("data/dat_scale.rds"))
dat_scale_ln <- readRDS(here::here("data/dat_scale_ln.rds"))

dat_sf <- dat_scale
dat_sf <- dat_sf %>%
  mutate(rate_prev_tbcases = (n_prev_tbcases / tent_cxr_total) * 100000)

dat_prev_sf <- dat_sf %>%
  mutate(
    name = "TB prevalence",
    tb_year = 2019,
    tb_year = as.factor(tb_year)
  ) %>%
  select(cluster, tb_year, name, value = rate_prev_tbcases, geometry)

dat_sf <- st_as_sf(dat_sf)
dat_sf$cluster <- factor(readr::parse_number(dat_scale$cluster))

# convert to sf object
dat_scale_ln_sf <- st_as_sf(dat_scale_ln)

dat_scale_ln_sf$cluster <- factor(readr::parse_number(dat_scale_ln_sf$cluster))

dat_scale_ln_sf <- dat_scale_ln_sf %>%
  mutate(rate_total_confirmed = (total_confirmed / population) * 100000)

dat_notif_prev_sf <- dat_scale_ln_sf %>%
  mutate(name = "TB notification") %>%
  select(cluster, tb_year, name, value = rate_total_confirmed)

dat_notif_prev_sf <- bind_rows(dat_notif_prev_sf, dat_prev_sf)

dat_notif_prev_sf$tb_year <- fct_relevel(
  as.character(dat_notif_prev_sf$tb_year),
  "2015", "2016", "2017", "2018", "2019"
)

dat_notif_prev_sf <- dat_notif_prev_sf %>%
  mutate(name = str_c(name, ": ", tb_year))
```


## 4. Baseline table comparing cluster vs. non cluster residents. TB notifications.

```{r tidy, results="asis"}
#tab_baseline <- tableby(resident ~ year + sex + age + tbclass + dotoption + hivstatus + any_micro_confirm,
#                        data = all_cases_sf %>% 
#                          mutate(year = factor(year)))
#summary(tab_baseline, title = "Baseline characteristics of patients with coordinates. Residents vs non cluster residents")
```

## 5. Table of cluster level statistics

```{r clustertable}
library(gtsummary) # produce summary tables

dat_sf <- dat_sf %>%
  mutate(n_tbcases = total_tbcases_2015 + total_tbcases_2016 +
    total_tbcases_2017 + total_tbcases_2018 + total_tbcases_2019)

dat_sf <- dat_sf %>%
  mutate(
    rate_tbcases_2015 = (total_tbcases_2015 / pop2015) * 100000,
    rate_tbcases_2016 = (total_tbcases_2016 / pop2016) * 100000,
    rate_tbcases_2017 = (total_tbcases_2017 / pop2017) * 100000,
    rate_tbcases_2018 = (total_tbcases_2018 / pop2018) * 100000,
    rate_tbcases_2019 = (total_tbcases_2019 / pop2019) * 100000,
    co_rate_tbcases_2015 = (n_confirmed_2015 / pop2015) * 100000,
    co_rate_tbcases_2016 = (n_confirmed_2016 / pop2016) * 100000,
    co_rate_tbcases_2017 = (n_confirmed_2017 / pop2017) * 100000,
    co_rate_tbcases_2018 = (n_confirmed_2018 / pop2018) * 100000,
    co_rate_tbcases_2019 = (n_confirmed_2019 / pop2019) * 100000,
    rate_prev_tbcases = (n_prev_tbcases / tent_cxr_total) * 100000,
    clinic_distance_km = clinic_distance / 1000,
    perce_maleadults = (male_adults / (male_adults + female_adults)) * 100,
    total_adults_n = male_adults + female_adults,
    perce_hiv_prev = hiv_prev * 100,
    perce_never_primary = never_primary * 100,
    never_primar_n = never_primary * intervadults
  )


dat_summary_table <- st_set_geometry(dat_sf, NULL) %>%
  select(
    rate_tbcases_2015,
    rate_tbcases_2016, rate_tbcases_2017,
    rate_tbcases_2018, rate_tbcases_2019,
    co_rate_tbcases_2015,
    co_rate_tbcases_2016, co_rate_tbcases_2017,
    co_rate_tbcases_2018, co_rate_tbcases_2019,
    rate_prev_tbcases, total_tbcases_2015, total_tbcases_2016,
    total_tbcases_2017, total_tbcases_2018, total_tbcases_2019,
    n_confirmed_2015, n_confirmed_2016,
    n_confirmed_2017, n_confirmed_2018, n_confirmed_2019,
    n_prev_tbcases,
    pop2015, pop2016, pop2017, pop2018, pop2019, tent_cxr_total, prop_adults,
    total_adults_n, perce_maleadults, male_adults,
    perce_never_primary, never_primar_n, intervadults,
    clinic_distance_km, perce_hiv_prev,
    hiv_tests_n, hiv_positive_n
  ) %>%
  tbl_summary(
    type = list(n_prev_tbcases ~ "continuous", n_confirmed_2019 ~ "continuous"),
    statistic = list(
      all_continuous() ~ " (mean: {mean}, range: {min}-{max}, sd:{sd})",
      c(
        total_tbcases_2015, total_tbcases_2016,
        total_tbcases_2017, total_tbcases_2018,
        total_tbcases_2019,
        n_confirmed_2015, n_confirmed_2016,
        n_confirmed_2017, n_confirmed_2018,
        n_confirmed_2019,
        n_prev_tbcases,
        pop2015, pop2016,
        pop2017, pop2018, pop2019, tent_cxr_total,
        never_primar_n, intervadults, total_adults_n, male_adults,
        # hiv_tests_n,n_hiv_positive
      ) ~ "(sum: {sum}, range: {min}-{max}, sd: {sd})"
    ),
    digits = list(all_continuous() ~ 1, c(perce_hiv_prev, clinic_distance_km, perce_maleadults) ~ 2),
    label = list(
      rate_tbcases_2015 ~ "2015 TB notification rate per 100000",
      rate_tbcases_2016 ~ "2016 TB notification rate per 100000",
      rate_tbcases_2017 ~ "2017 TB notification rate per 100000",
      rate_tbcases_2018 ~ "2018 TB notification rate per 100000",
      rate_tbcases_2019 ~ "2019 TB notification rate per 100000",
      co_rate_tbcases_2015 ~ "2015 confirmed TB notification rate per 100000",
      co_rate_tbcases_2016 ~ "2016 confirmed TB notification rate per 100000",
      co_rate_tbcases_2017 ~ "2017 confirmed TB notification rate per 100000",
      co_rate_tbcases_2018 ~ "2018 confirmed TB notification rate per 100000",
      co_rate_tbcases_2019 ~ "2019 confirmed TB notification rate per 100000",
      rate_prev_tbcases ~ "2019 TB prevalence rate per 100000",
      total_tbcases_2015 ~ "2015 TB notified cases",
      total_tbcases_2016 ~ "2016 TB notified cases",
      total_tbcases_2017 ~ "2017 TB notified cases",
      total_tbcases_2018 ~ "2018 TB notified cases",
      total_tbcases_2019 ~ "2019 TB notified cases",
      n_confirmed_2015 ~ "2015 confirmed TB notified cases",
      n_confirmed_2016 ~ "2016 confirmed TB notified cases",
      n_confirmed_2017 ~ "2017 confirmed TB notified cases",
      n_confirmed_2018 ~ "2018 confirmed TB notified cases",
      n_confirmed_2019 ~ "2019 confirmed TB notified cases",
      n_prev_tbcases ~ "2019 TB prevalent cases",
      pop2015 ~ "2015 population",
      pop2016 ~ "2016 population",
      pop2017 ~ "2017 population",
      pop2018 ~ "2018 population",
      pop2019 ~ "2019 population",
      tent_cxr_total ~ "2019 prevalence survey population",
      prop_adults ~ "Percentage of adults (>=15y) (%)",
      total_adults_n ~ "Adults (>=15y) (n)",
      perce_maleadults ~ "Percentage of male adults (%)",
      male_adults ~ "Male adults (n)",
      clinic_distance_km ~ "Distance to TB clinic (km)",
      perce_never_primary ~ "Household head without primary education (%)",
      intervadults ~ "Prev interviewed (N)",
      never_primar_n ~ "Household head without primary education (n)",
      perce_hiv_prev ~ "HIV prevalence (%)",
      hiv_tests_n ~ "Total HIV tests (n)",
      hiv_positive_n ~ "HIV prevalence (n)"
    ),
    missing_text = "(Missing)"
  ) %>%
  as_flex_table()

save_as_docx(
  "Cluster level summary data" = dat_summary_table,
  path = here::here("figures/table1.docx")
)

dat_summary_table
```

## 6. Graphs of all the variables
```{r all_predictors}
# Graphs of all the variables below
g <- dat_sf %>%
  select(cluster, geometry)

dat_sf_vars <- dat_sf %>%
  mutate(
    never_primary = never_primary * 100,
    hiv_prev = hiv_prev * 100,
    clinic_distance = clinic_distance / 1000
  ) %>%
  select(cluster,
    "(A) Total Population 2015" = total,
    "(B) Percent adults (%) (>=15y)" = prop_adults,
    "(C) Percent male adults (%)" = perc_male,
    "(D) Distance to nearest\n      TB clinic (km)" = clinic_distance,
    "(E) HIV prevalence (%)" = hiv_prev,
    "(F) Household head without \n      primary education (%)" = never_primary
  ) %>%
  st_set_geometry(NULL) %>%
  pivot_longer(-cluster) %>%
  left_join(g, by = "cluster") %>%
  st_as_sf()

# library(cowplot)
# dat_sf_vars %>%
#     split(.$name) %>%
#   map(~ggplot(.,aes(fill = value)) +
#   geom_sf() +
#     theme_minimal() +
#     scale_fill_viridis(option = "D", direction = -1, name='')+
#      coord_sf(expand = F) +
#   scale_x_continuous(breaks = c(34.98, 35.02, 35.06, 35.10)) +
#   scale_y_continuous(breaks = c(-15.86, -15.82, -15.78, -15.74)) +
#   theme(axis.text.x=element_text(angle=45, hjust=1)) +
#   facet_wrap(~name)) %>%
#   cowplot::plot_grid(plotlist = .)




dat_sf_vars_df <- dat_sf_vars %>%
  group_nest(name)

plot_sf <- function(df_sf, title_sf) {
  ggplot(df_sf, aes(fill = value)) +
    geom_sf() +
    theme_minimal() +
    scale_fill_viridis(option = "D", name = "") +
    coord_sf(expand = F) +
    scale_x_continuous(breaks = c(34.98, 35.02, 35.06, 35.10)) +
    scale_y_continuous(breaks = c(-15.86, -15.82, -15.78, -15.74)) +
    labs(title = title_sf) +
    ggspatial::annotation_scale(pad_x = unit(0.25, "cm"),
  pad_y = unit(0.15, "cm")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title.position = "plot",
      plot.title = element_text(size = 11)
    ) # make the title align with yaxis labels
}

dat_sf_vars_df <- dat_sf_vars_df %>%
  mutate(plots = map2(data, name, plot_sf))

patchwork_dat_sf <- dat_sf_vars_df$plots[[1]] + dat_sf_vars_df$plots[[2]] + dat_sf_vars_df$plots[[3]] +
  dat_sf_vars_df$plots[[4]] + dat_sf_vars_df$plots[[5]] + dat_sf_vars_df$plots[[6]]

patchwork_dat_sf

ggsave(here::here("figures/Fig1.tiff"), width = 25, height = 15, units = "cm")
```
## 7. All TB rates
```{r all_TB_rates}
dat_notif_prev_df <- dat_notif_prev_sf %>%
  group_nest(name)

initials <- str_c("(", LETTERS[1:6], ")")
initials

dat_notif_prev_df <- dat_notif_prev_df %>%
  mutate(name = str_c(initials, " ", name))

# Find the min and max of notifications and prevalances
dat_notif_prev_df %>%
  mutate(min_rate = map_dbl(data, ~ min(.x[["value"]])), max_rate = map_dbl(data, ~ max(.x[["value"]])))

plot_sfnot <- function(df_sf, title_sf) {
  ggplot(df_sf, aes(fill = value)) +
    geom_sf() +
    theme_minimal() +
    scale_fill_viridis(option = "D", name = "", limits = c(0, 400)) +
    coord_sf(expand = F) +
    scale_x_continuous(breaks = c(34.98, 35.02, 35.06, 35.10)) +
    scale_y_continuous(breaks = c(-15.86, -15.82, -15.78, -15.74)) +
    labs(title = title_sf) +
    ggspatial::annotation_scale(pad_x = unit(0.25, "cm"),
  pad_y = unit(0.15, "cm")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title.position = "plot",
      plot.title = element_text(size = 11)
    ) # make the title align with yaxis labels
}

plot_sfprev <- function(df_sf, title_sf) {
  ggplot(df_sf, aes(fill = value)) +
    geom_sf() +
    theme_minimal() +
    scale_fill_viridis(option = "E", name = "", limits = c(0, 1500)) +
    coord_sf(expand = F) +
    scale_x_continuous(breaks = c(34.98, 35.02, 35.06, 35.10)) +
    scale_y_continuous(breaks = c(-15.86, -15.82, -15.78, -15.74)) +
    labs(title = title_sf) +
    ggspatial::annotation_scale(pad_x = unit(0.25, "cm"),
  pad_y = unit(0.15, "cm")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title.position = "plot",
      plot.title = element_text(size = 11)
    ) # make the title align with yaxis labels
}

dat_notif_plots <- dat_notif_prev_df %>%
  filter(name != "(F) TB prevalence: 2019") %>%
  mutate(name = str_replace(name, pattern = "TB notification", replacement = "TB CNR")) %>%
  mutate(plots = map2(data, name, plot_sfnot))


dat_prev_plots <- dat_notif_prev_df %>%
  filter(name == "(F) TB prevalence: 2019") %>%
  mutate(name = str_replace(name, pattern = "TB prevalence", replacement = "TB CPR")) %>%
  mutate(plots = map2(data, name, plot_sfprev))

notif_prev_plots <- dat_notif_plots$plots[[1]] + dat_notif_plots$plots[[2]] + dat_notif_plots$plots[[3]] +
  dat_notif_plots$plots[[4]] + dat_notif_plots$plots[[5]] + dat_prev_plots$plots[[1]]

notif_prev_plots

ggsave(here::here("figures/Fig2.tiff"), width = 25, height = 15, units = "cm")
```

<br>

## 8. Reproducibility

This reproduction of the analysis was run by:

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName=names(sysinfo), value=sysinfo, row.names=NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
