---
title: "TBhotspots"

subtitle: "Comparing denominators derived based on two methods, one WorldPOP other SCALE census"

author: |
  | McEwen Khundi
  |
  | London School of Tropical Hygiene & Medicine, London, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    df_print: paged
    theme: spacelab
    toc: true
---

<br>

## 1. Backgound


<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, message=FALSE}
library(tidyverse) # for data manipulation
library(pmthemes) # for ggplot themes
library(knitr) # for tables
library(sf)
library(viridis)
```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import}
load(here::here("data/all_cases_sf.rda"))
load(here::here("data/scale_clusters_sf.rda"))

# Population denominators estimated from the world pop data start with wp
# From scale census
load(here::here("data/wp_scale_cluster_pops_2015to2019.rda"))
load(here::here("data/scalebtcity_census_2015_2019_cluster.rda"))
```

<br>

## 4. Tidying and recoding

```{r tidy}
cluster_pops_wp <- wp_scale_cluster_pops_2015to2019 %>%
  group_by(c2, year) %>%
  summarise(wp_population = sum(pop), .groups = "drop") %>%
  mutate(cluster = paste0("c", c2)) %>%
  select(-c2)

head(cluster_pops_wp, 20)

cluster_pops_sc <- scalebtcity_census_2015_2019_cluster %>%
  group_by(cluster, year) %>%
  summarise(sc_population = sum(population), .groups = "drop") %>%
  mutate(cluster = paste0("c", readr::parse_number(cluster)), )


head(cluster_pops_sc, 20)

cluster_pops <- full_join(cluster_pops_wp, cluster_pops_sc,
  by = c("cluster", "year")
) %>%
  select(cluster, year, wp_population, sc_population)

cluster_notifications <- all_cases_sf %>%
  st_set_geometry(NULL) %>%
  group_by(cluster, year) %>%
  summarise(notifs = n(), .groups = "drop") %>%
  filter(!is.na(cluster), between(year, 2015, 2019))

head(cluster_notifications, 20)


cnr_2015to2019 <- full_join(cluster_pops,
  cluster_notifications,
  by = c("year", "cluster")
) %>%
  relocate(c(
    "cluster", "year", "notifs",
    "wp_population", "sc_population"
  ))

head(cnr_2015to2019, 20)
```

<br>

## 5. Comparing the world pop versus the scale interpolated denominators
```{r}
cluster_pops %>%
  ggplot(aes(x = wp_population, y = sc_population)) +
  geom_point() +
  facet_wrap(~year)
```


<br>

## 6. Cases notification rates calculations
Using the world population cluster denominators and the SCALE census interpolated denominators

```{r cnrs}
# Need to calculate the confidence intervals
cnr_2015to2019 <- cnr_2015to2019 %>%
  mutate(
    wp_cnr = notifs / wp_population,
    wp_cnr_100 = wp_cnr * 100000
  ) %>%
  mutate(
    sc_cnr = notifs / sc_population,
    sc_cnr_100 = sc_cnr * 100000
  )


cnr_2015to2019 <- full_join(cnr_2015to2019,
  scale_clusters_sf,
  by = "cluster"
) %>%
  st_as_sf() %>%
  mutate(cluster = factor(cluster, levels = paste0("c", 1:72)))

cnr_2015to2019_df <- cnr_2015to2019 %>%
  st_set_geometry(NULL) %>%
  mutate(year = factor(year, levels = 2015:2019))

head(cnr_2015to2019_df, 30)
```

<br>

## 7. Case notification rates graphs per cluster based on world population denominators

```{r wp_cnrs_graphs}


# Cluster line graphs, example cluster one
cnr_2015to2019_df %>%
  filter(cluster %in% c(paste0("c", 1))) %>%
  ggplot(aes(x = year, y = wp_cnr_100)) +
  geom_line(aes(group = 1)) +
  geom_point(aes(color = year, group = year)) +
  labs(title = "WP TB notification rates per cluster per 100K population") +
  theme_bw()


# Cluster line graphs
cnr_2015to2019_df %>%
  # filter(cluster %in% c(paste0("c",1:20)))%>%
  ggplot(aes(x = year, y = wp_cnr_100)) +
  geom_line(aes(group = 1)) +
  geom_point(aes(color = year, group = year)) +
  labs(title = "WP TB notification rates per cluster per 100K population") +
  theme_bw() +
  facet_wrap(~cluster, ncol = 12)

# Cluster map based
cnr_2015to2019 %>%
  ggplot(mapping = aes(fill = wp_cnr_100)) +
  geom_sf() +
  scale_fill_viridis(
    option = "magma", direction = -1,
    name = "CNR per 100"
  ) +
  labs(title = "WP TB notification rates per cluster per 100K population") +
  theme_bw() +
  facet_wrap(~year)
```

## 8. Case notification rates graphs per cluster based on SCALE census interpolated denominators

```{r sc_cnrs_graphs}

# Cluster line graphs, example cluster one
cnr_2015to2019_df %>%
  filter(cluster %in% c(paste0("c", 1))) %>%
  ggplot(aes(x = year, y = sc_cnr_100)) +
  geom_line(aes(group = 1)) +
  geom_point(aes(color = year, group = year)) +
  labs(title = "SC TB notification rates per cluster per 100K population") +
  theme_bw()


# Cluster line graphs
cnr_2015to2019_df %>%
  # filter(cluster %in% c(paste0("c",1:20)))%>%
  ggplot(aes(x = year, y = sc_cnr_100)) +
  geom_line(aes(group = 1)) +
  geom_point(aes(color = year, group = year)) +
  labs(title = "SC TB notification rates per cluster per 100K population") +
  theme_bw() +
  facet_wrap(~cluster, ncol = 12)

# Cluster map based
cnr_2015to2019 %>%
  ggplot(mapping = aes(fill = sc_cnr_100)) +
  geom_sf() +
  scale_fill_viridis(
    option = "magma", direction = -1,
    name = "CNR per 100"
  ) +
  labs(title = "SC TB notification rates per cluster per 100K population") +
  theme_bw() +
  facet_wrap(~year)
```
<br>

## 9. Case notification rates graphs per cluster compared between world population and the scale

```{r wp__sc_cnrs_graphs}


# Cluster line graphs, example cluster one
cnr_2015to2019_df %>%
  filter(cluster %in% c(paste0("c", 1))) %>%
  ggplot(aes(x = sc_cnr_100, y = wp_cnr_100)) +
  geom_line(aes(group = 1)) +
  geom_point(aes(color = year, group = year)) +
  labs(title = "SC vs WP TB notification rates per cluster per 100K population") +
  theme_bw()


# Cluster line graphs
cnr_2015to2019_df %>%
  # filter(cluster %in% c(paste0("c",1:20)))%>%
  ggplot(aes(x = sc_cnr_100, y = wp_cnr_100)) +
  geom_line(aes(group = 1)) +
  geom_point(aes(color = year, group = year)) +
  labs(title = "SC vs WP TB notification rates per cluster per 100K population") +
  theme_bw() +
  facet_wrap(~cluster, ncol = 12)
```
<br>

## 10. Reproducibility

This reproduction of the analysis was run by: 

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName = names(sysinfo), value = sysinfo, row.names = NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
